CF Nutrition Pre- and Post-CFTR Modulator

### 2022-09-30 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
Update cf_nutrition.qmd

---
title: "CF Nutrition Pre- and Post-CFTR Modulator"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    self-contained: true
editor: visual
---

@@

```{r}

@@

vit <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
mods <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
pfts <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
lfts <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_LFTs_ThroughDecember2021.xlsx")
pi = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Fecal Elastase_throughDecember2021.xlsx")

@@

pi = pi %>% select(`Patient ID`,Date,`Fecal Elastase`)
pi$`Fecal Elastase` = as.numeric(pi$`Fecal Elastase`)
pi$Date = as.Date(pi$Date)

@@

vit = full_join(vit,pi,by = c("Patient ID", "Date"))

@@

            across(`Days from Start`:`Fecal Elastase`, ~ mean(.x, na.rm = TRUE)))

@@

            across(`Days from Start`:`Fecal Elastase`,

@@

df$`Pancreatic Insufficient` = factor(df$`Fecal Elastase` < 128,

@@

@@

@@

-   Pancreatic insufficiency was determined based on a fecal elastase cutoff of \< 128 $\mu g/g$.

@@

# Baseline value grouped by first modulator

```{r results='asis'}
t1 <- tableby(`First Modulator` ~ `Age at Start` + Sex + `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)` + `Fecal Elastase` + `Pancreatic Insufficient`,data = df[df$Timepoint == "Pre", ],numeric.stats = c("Nmiss","medianrange"))
summary(t1, pfootnote = T)
```

# Pre/post comparison

```{r results='asis'}
p <- paired(Timepoint ~ `VitaminA_Retinol` + signed.rank(`25OH-Vitamin D`,"Nmiss","median","q1q3") + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)` + `Pancreatic Insufficient`,data = df, id = `Patient ID`)
summary(p, pfootnote = T)
```

>>>>

---
title: "CF Nutrition Pre- and Post-CFTR Modulator"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
editor: visual
---

@@

```{r warning=FALSE}

@@

vit <- read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
mods <- read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
pfts <- read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
lfts <- read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_LFTs_ThroughDecember2021.xlsx")
geno = read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/CF Info/Sweat Tests from Internal Database_genotype class_withPancStatus.xlsx",na = ".")

@@

geno = geno %>% select(PatientID,genoRisk,PancreaticStatus) %>% distinct() %>% 
  rename(`Patient ID`=PatientID,`Genotype Risk` = genoRisk,
         `Pancreatic Status`=PancreaticStatus)

@@

@@

            across(`Days from Start`:GGTP, ~ mean(.x, na.rm = TRUE)))

@@

            across(`Days from Start`:GGTP,

@@

@@

df$`Elevated LFTs (3X)`[rowSums(is.na(df[,c("ALT","AST","GGTP")]))==3]=NA

@@

df$`Elevated LFTs (5X)`[rowSums(is.na(df[,c("ALT","AST","GGTP")]))==3]=NA
# Add genetic info
df = left_join(df,geno,by = "Patient ID")
# Delete those without paired obs.
df = df %>%  group_by(`Patient ID`) %>% filter(n() == 2) %>% 
  arrange(`Patient ID`,Timepoint)
# Write data so Edith can check
write.csv(df,file = "/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_data.csv",row.names = F,na="")
write.csv(df[df$`First Modulator`=="Trikafta",],file = "/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_data_eti_only.csv",row.names = F,na="")

@@

@@

# Ivacaftor and ETI

## Baseline value grouped by first modulator

```{r results='asis'}
t1 <- tableby(`First Modulator` ~ `Age at Start` + Sex + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)`+`Pancreatic Status`,data = df[df$Timepoint == "Pre", ],numeric.stats = c("Nmiss","medianrange"),numeric.test="kwt")
summary(t1, pfootnote = T)
```

## Pre/post comparison

```{r results='asis'}
p <- paired(Timepoint ~ `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)`,data = df, id = `Patient ID`,numeric.stats = c("Nmiss","medianrange"),numeric.test="signed.rank",signed.rank.exact = FALSE)
summary(p, pfootnote = T)
```

## Pre/post comparison by pancreatic status

One participant with unknown pancreatic status was removed in order to calculate p values.

```{r results='asis'}
panc_df = df %>% filter(`Pancreatic Status`!= "Unknown")

p <- paired(Timepoint ~ `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)`,data = panc_df, id = `Patient ID`,strata = `Pancreatic Status`,numeric.stats = c("Nmiss","medianrange"),numeric.test="signed.rank",signed.rank.exact = FALSE)
summary(p, pfootnote = T)
```

# ETI only

## Pre/post comparison

```{r results='asis'}
p <- paired(Timepoint ~ `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)` + `Pancreatic Status`,data = df[df$`First Modulator`=="Trikafta",], id = `Patient ID`,numeric.stats = c("Nmiss","medianrange"),numeric.test="signed.rank",signed.rank.exact = FALSE)
summary(p, pfootnote = T)
```
